---
title: "miRNAcounts"
output: html_document
date: '2023-07-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
library(wesanderson)
library("RColorBrewer")
library(ggrepel)
```

These bash and R commands are aiming to get the miRNA counts in libraries to see the correlation between the two Piwi6KO lines. 
To normalize, to the number of library reads passing the filters (filter for minimum length of 15, max length 35) after adapter trimming were used. 
```{bash}
ezgit@octarine:~/WD_202010/miRNAcounts_202307$ cat countingmiRNAs.sh
#For this script, ONLY genome mapped .bed files of Piwi6KO#10 (RDVJ189.bed) and Piwi6KO#52 (RDVJ190.bed) should be in the directory.
#count the numver of miRNAs and normalize to reads mapping filter and use it for the correlation
WORKDIR=`pwd`
for files in "${WORKDIR}"/*.bed; do
        FILENAME=$(echo $(basename "${files}") | sed 's/\.bed$//')
        BED="${FILENAME}".bed
        MIRNAS="${FILENAME}".miRNAs

bedtools coverage -b "${BED}" -a $HOME/U4.4_microRNAs.bed -c | awk 'BEGIN{OFS="\t"} \
{print$1,$2,$3,$4,$5,$6,$7+1}'  > "${MIRNAS}"
done

#nrm to reads passing filters
#filtered reads: 17523988 for Piwi6KO#10 and 14919801 for Piwi6KO#52. 
paste -d "\t" *.miRNAs | awk  -v OFS="\t" '{print$1,$2,$3,$4,$5,$6,$7/17523988*1000000,$14/14919801*1000000}' > nrmtoreadspassing_filters.txt

cat nrmtoreadspassing_filters.txt | awk  -v OFS="\t" '{print$1,$2,$3,$4,$5,$6,log($7)/log(2),log($8)/log(2)}' > nrmtoreadspassing_filterslog2.txt

```


Figure s3A
Save nrmtoreadspassing_filterslog2.txt to outputfrombash in the directory where you want to run the R commands. 
```{r}
miRNAcounts <- read_table("./outputfrombash/nrmtoreadspassing_filterslog2.txt",col_names=c("chromosome","start","end","miRNA",".","strand","PIWI6KO#10","PIWI6KO#52"), col_types = c(col_character(),col_number(),col_number(),col_character(),col_character(),col_character(),col_number(), col_number()))
miRNAcounts <- as.data.frame(miRNAcounts)


plotmiRNAs <- ggplot(miRNAcounts, aes(x=miRNAcounts$`PIWI6KO#10`, y=miRNAcounts$`PIWI6KO#52`)) + geom_point(size=6, alpha = .6,color="#0A8974") +     theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())  + geom_abline(slope=1,linetype="dashed", color="black")+ xlim(-5,15)+ylim(-5,15)


png(file="plotmiRNAs.png",
    width=382, height=484)
plotmiRNAs
dev.off()

```

