---
title: "TE-quantification-SalmonMapping_Piwi6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(DESeq2)
library(tximport)
library(tximport)
library(ggforce)
library(clusterProfiler)
library(latticeExtra)
library(rtracklayer)
library(ggplot2)
library(ggrepel)
library(pheatmap)
library(RColorBrewer)
library("ggvenn")
library("ggVennDiagram")
```


Condition as design. 
Metadata3.csv
Library	Name	Batch	Clonal	Clone	Genotype	Condition
RDVJ640	PIWI5KO_O_1	1	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ641	PIWI5KO_O_2	2	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ642	PIWI5KO_O_3	3	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ643	PIWI5KO_A_1	1	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ644	PIWI5KO_A_2	2	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ645	PIWI5KO_A_3	3	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ646	PIWI6KO_10_1	1	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ647	PIWI6KO_10_2	2	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ648	PIWI6KO_10_3	3	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ649	PIWI6KO_52_1	1	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ650	PIWI6KO_52_2	2	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ651	PIWI6KO_52_3	3	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ652	CTRLKO_1_1	1	yes	clone1	control_1	control
RDVJ653	CTRLKO_1_2	2	yes	clone1	control_1	control
RDVJ654	CTRLKO_1_3	3	yes	clone1	control_1	control
RDVJ655	CTRLKO_6_1	1	yes	clone6	control_6	control
RDVJ656	CTRLKO_6_2	2	yes	clone6	control_6	control
RDVJ657	CTRLKO_6_3	3	yes	clone6	control_6	control
RDVJ658	WT_1	1	no	clone0	WT	WT
RDVJ659	WT_2	2	no	clone0	WT	WT
RDVJ660	WT_3	3	no	clone0	WT	WT

```{r}
#Metadata
metadata <- read_delim("Metadata/Metadata3.csv", delim = ";",col_names = TRUE,col_types = c(col_character(),col_character(),col_double(),col_character(),col_double(),col_character(),col_character()))%>% arrange(Library)
```

Condition as the design, and take control as the baseline. 
```{r}
SampleTable <- metadata %>%
  mutate(Condition= as.factor(Condition)) %>%
  arrange(Library) %>%
  as.data.frame() %>%
  mutate(Condition = fct_relevel(Condition, "control"))
         
SampleNames = c("PIWI5KO_O_1", "PIWI5KO_O_2", "PIWI5KO_O_3", "PIWI5KO_A_1", "PIWI5KO_A_2", "PIWI5KO_A_3","PIWI6KO_10_1", "PIWI6KO_10_2", "PIWI6KO_10_3", "PIWI6KO_52_1", "PIWI6KO_52_2", "PIWI6KO_52_3","CTRLKO_1_1", "CTRLKO_1_2", "CTRLKO_1_3","CTRLKO_6_1", "CTRLKO_6_2", "CTRLKO_6_3", "WT_1","WT_2","WT_3")

Matrix_TEs = read_delim("Metadata/TEsLau_transcript",
                        delim = "\t",
                        col_names = FALSE)     
```

Takes Salmon quantification data as input
```{r}
#Salmon quantification data

Dir  = "CountDataTE"
FilePath = dir("CountDataTE/", ".sf")
Files=file.path(Dir, FilePath)

names(Files) = paste0(SampleNames)

QuantFiles = tximport(Files, type = "salmon",  tx2gene = Matrix_TEs)
```


General Settings
```{r}
#p value for DESeq2 analysis
pvalue <- 0.01
#log2 fold change for DESeq2 analysis
lfc <- 1
#minimal count of TE to be taken along in statistical analysis
lowCounts <- 10
#FDR
FDR <- 0.05
```



```{r}
#Create a DESeqDataSet
ddsTE = DESeqDataSetFromTximport(QuantFiles,SampleTable,~Condition)
colnames(ddsTE)
#Set the reference as control
ddsTE$Condition <- relevel(ddsTE$Condition, ref = "control")


#Check PCA plots
rld <- rlog(ddsTE, blind =FALSE)
PCAcondition <- plotPCA(rld, intgroup= "Condition")
print(PCAcondition)
png(file="PCAcondition.png",
   width=429, height=484)
PCAcondition
dev.off()

pdf(file="PCAcondition.pdf")
PCAcondition
dev.off()

PCAgenotype_condition <- plotPCA(rld, intgroup= c("Genotype","Condition"))
print(PCAgenotype_condition)
png(file="PCAgenotype_condition.png",
   width=429, height=484)
PCAgenotype_condition
dev.off()
pdf(file="PCAgenotype_condition.pdf")
PCAgenotype_condition
dev.off()



#filter low counts
ddsTE = ddsTE[rowSums(counts(ddsTE)) > lowCounts, ]

#run DEseq
DEseq_TE = DESeq(ddsTE,quiet = TRUE)
resultsNames(DEseq_TE)
#[1] "Intercept"                    "Condition_PIWI5KO_vs_control"
#[3] "Condition_PIWI6KO_vs_control" "Condition_WT_vs_control"
```

Cooks distance

```{r}
boxplot(log10(assays(DEseq_TE)[["cooks"]]), cex=0.5)
```


plot sample distance
```{r}
vst <- vst(DEseq_TE, blind = TRUE)
sampleDists <- dist(t(assay(vst)))
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) = vst$Name
colnames(sampleDistMatrix) = NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors,
         main = "Sample distance")
```

Dispersion
```{r}
plotDispEsts(DEseq_TE)
```


"Condition_PIWI6KO_vs_control"
Comparing PIWI6KO to control lines.
```{r}
#Piwi6KO vs control 
results_TE_Piwi6vscontrol <- results(DEseq_TE,alpha= pvalue,contrast=c("Condition", "PIWI6KO", "control"))
plotMA(results_TE_Piwi6vscontrol,ylim=c(-10,10))
summary(results_TE_Piwi6vscontrol)
results_TE_Piwi6vscontrol <- results_TE_Piwi6vscontrol %>% as.data.frame() %>% rownames_to_column(var = "TE") %>% as_tibble()
#TEs significantly upregulated at least 2-fold
filter(results_TE_Piwi6vscontrol, padj <= pvalue & log2FoldChange >= lfc)


counts_TEs_Piwi6vscontrol = counts(DEseq_TE,normalized=TRUE,replaced=FALSE) %>%
 as.data.frame() %>%
 rownames_to_column(var = "TE") %>%
 as_tibble %>%
 group_by(TE) %>%
summarize(meanWT=mean(c(WT_1,WT_2,WT_3)),
 meanPIWI5KO=mean(c(PIWI5KO_O_1,PIWI5KO_O_2,PIWI5KO_O_3,PIWI5KO_A_1,PIWI5KO_A_2,PIWI5KO_A_3)),
 meanPIWI6KO=mean(c(PIWI6KO_10_1,PIWI6KO_10_2,PIWI6KO_10_3,PIWI6KO_52_1,PIWI6KO_52_2,PIWI6KO_52_3)),
 meanCTRLKO=mean(c(CTRLKO_1_1,CTRLKO_1_2,CTRLKO_1_3,CTRLKO_6_1,CTRLKO_6_2,CTRLKO_6_3)))%>%
 mutate(meanWT= meanWT + 0.5,
 meanPIWI5KO= meanPIWI5KO + 0.5,
 meanPIWI6KO= meanPIWI6KO + 0.5,
 meanCTRLKO=meanCTRLKO + 0.5) %>%
 left_join(results_TE_Piwi6vscontrol, by = "TE") %>%
 mutate(padj = replace_na(padj, 1),
 sign = ifelse(padj <= FDR, "significant", "NS"), 
 TEtypeSign = str_extract(TE, "(?<=\\d@)\\w+"))
 #TEtypeSign = fct_relevel(TEtypeSign, c("LTR_Gypsy", "DNA_hAT", "RC_Helitron", "Unknown")))

#To categorize the TEs according to their significant changes upon Piwi6KO.
write.table(counts_TEs_Piwi6vscontrol, file = "counts_TEs_Piwi6vscontrol.tsv", row.names=FALSE, sep="\t")
nonsignificant <- counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign=="NS",1]
write.table(nonsignificant, file = "nonsignificant.tsv", row.names=FALSE, sep="\t")
 significant_downregulated <- counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign=="significant" &counts_TEs_Piwi6vscontrol$log2FoldChange<0,1]
write.table(significant_downregulated, file = "significant_downregulated.tsv", row.names=FALSE, sep="\t")
significant_upregulated <- counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign=="significant" &counts_TEs_Piwi6vscontrol$log2FoldChange>0,1]
write.table(significant_upregulated, file = "significant_upregulated.tsv", row.names=FALSE, sep="\t")
significant_upregulatedmorethan4 <-  counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign=="significant" &counts_TEs_Piwi6vscontrol$log2FoldChange>4,1]
write.table(significant_upregulatedmorethan4, file = "significant_upregulatedmorethan4.tsv", row.names=FALSE, sep="\t")


#Figure 3A, TEs colored
plotPiwi6vscontrolcolors <- ggplot()+ geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "NS", ], aes(meanCTRLKO, meanPIWI6KO),size = 6,alpha=0.6,colour = "gray")+ geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR") , ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#0A8974")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "Unknown" , ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#C2C8E7")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I") , ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E27520")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator") , ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E49C23")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "RC_Helitron", ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E3072A")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "Satellite", ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#6362A9")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "SINE_tRNA", ], aes(meanCTRLKO, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#5FAECC")+
    scale_x_continuous(trans="log2", limits=c(0.5,NA)) + scale_y_continuous(trans="log2", limits=c(0.5,NA)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_abline(slope = 1, intercept= 1) +
    geom_abline(slope = 1, intercept= -1) +
theme(legend.title = element_blank(), panel.background = element_blank(), axis.line = element_line(),legend.key = element_blank()) +theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())

png(file="plotPiwi6vscontrolcolors_TE.png",
   width=484, height=484)
plotPiwi6vscontrolcolors
dev.off()

#volcano plot, #Figure 3B, TEs colored
volcanoplotPiwi6vscontrolwithouttextcolors <- ggplot()+ geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "NS", ], aes(log2FoldChange, -log10(padj)),size = 6,alpha=0.6,colour = "gray")+  
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#0A8974")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "Unknown" , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#C2C8E7")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E27520")+geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign %in% c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E49C23")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "RC_Helitron", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E3072A")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "Satellite", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#6362A9")+
geom_point(data = counts_TEs_Piwi6vscontrol[counts_TEs_Piwi6vscontrol$sign == "significant" & counts_TEs_Piwi6vscontrol$TEtypeSign == "SINE_tRNA", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#5FAECC")+xlim(-10,10)+theme(legend.title = element_blank(), panel.background = element_blank(), axis.line = element_line(),legend.key = element_blank())+geom_vline(xintercept=c(-2, 2), col="black",linetype="longdash")+ geom_hline(yintercept=-log10(0.05), col="black",linetype="longdash")+theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())

png(file="volcanoplotPiwi6vscontrolwithouttextcolors_TE.png",
   width=484, height=484)
volcanoplotPiwi6vscontrolwithouttextcolors
dev.off() 
```

Scatter plots to compare the two Piwi6KO clones: PIWI6KO10 vs PIWI6KO52
```{r}
counts_TEs_Piwi610and52 = counts(DEseq_TE,normalized=TRUE,replaced=FALSE) %>%
 as.data.frame() %>%
 rownames_to_column(var = "TE") %>%
 as_tibble %>%
 group_by(TE) %>%
 summarize(meanWT=mean(c(WT_1,WT_2,WT_3)),
 meanPIWI6KO_10=mean(c(PIWI6KO_10_1,PIWI6KO_10_2,PIWI6KO_10_3)),
 meanPIWI6KO_52=mean(c(PIWI6KO_52_1,PIWI6KO_52_2,PIWI6KO_52_3)),
 meanCTRLKO_1=mean(c(CTRLKO_1_1,CTRLKO_1_2,CTRLKO_1_3)),
 meanCTRLKO_6=mean(c(CTRLKO_6_1,CTRLKO_6_2,CTRLKO_6_3))) %>%
 mutate(meanWT= meanWT + 0.5,
 meanPIWI6KO_10= meanPIWI6KO_10 + 0.5,
 meanPIWI6KO_52= meanPIWI6KO_52 + 0.5,
 meanCTRLKO_1=meanCTRLKO_1 + 0.5,
 meanCTRLKO_6=meanCTRLKO_6 + 0.5) %>%
 left_join(results_TE_Piwi6vscontrol, by = "TE") %>%
 mutate(padj = replace_na(padj, 1),
 sign = ifelse(padj <= FDR, "significant", "NS"), 
 TEtypeSign = str_extract(TE, "(?<=\\d@)\\w+"))

#TEs colored, Figure S3D
scatterplotPiwi610vs52colors <- ggplot()+ geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "NS", ], aes(meanPIWI6KO_10, meanPIWI6KO_52),size = 6,alpha=0.6,colour = "gray")+ geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign %in% c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR") , ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#0A8974")+geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign == "Unknown" , ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#C2C8E7")+geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign %in% c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I") , ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E27520")+geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign %in% c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator") , ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E49C23")+
geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign == "RC_Helitron", ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E3072A")+
geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign == "Satellite", ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#6362A9")+
geom_point(data = counts_TEs_Piwi610and52[counts_TEs_Piwi610and52$sign == "significant" & counts_TEs_Piwi610and52$TEtypeSign == "SINE_tRNA", ], aes(meanPIWI6KO_10, meanPIWI6KO_52, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#5FAECC")+
    scale_x_continuous(trans="log2", limits=c(0.5,NA)) + scale_y_continuous(trans="log2", limits=c(0.5,NA)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_abline(slope = 1, intercept= 1) +
    geom_abline(slope = 1, intercept= -1) +
theme(legend.title = element_blank(), panel.background = element_blank(), axis.line = element_line(),legend.key = element_blank()) +theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())

png(file="scatterplotPiwi610vs52colors_TE.png",
   width=484, height=484)
scatterplotPiwi610vs52colors
dev.off()  
```