---
title: "TEquantification with WT as a base for comparison"
output: html_document
date: '2024-09-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
library(DESeq2)
library(tximport)
library(tximport)
library(ggforce)
library(clusterProfiler)
library(latticeExtra)
library(rtracklayer)
library(ggplot2)
library(ggrepel)
library(pheatmap)
library(RColorBrewer)
library("ggvenn")
library("ggVennDiagram")
```


Condition as design. 
Metadata3.csv
Library	Name	Batch	Clonal	Clone	Genotype	Condition
RDVJ640	PIWI5KO_O_1	1	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ641	PIWI5KO_O_2	2	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ642	PIWI5KO_O_3	3	yes	cloneO	PIWI5KO_O	PIWI5KO
RDVJ643	PIWI5KO_A_1	1	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ644	PIWI5KO_A_2	2	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ645	PIWI5KO_A_3	3	yes	cloneA	PIWI5KO_A	PIWI5KO
RDVJ646	PIWI6KO_10_1	1	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ647	PIWI6KO_10_2	2	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ648	PIWI6KO_10_3	3	yes	clone10	PIWI6KO_10	PIWI6KO
RDVJ649	PIWI6KO_52_1	1	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ650	PIWI6KO_52_2	2	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ651	PIWI6KO_52_3	3	yes	clone52	PIWI6KO_52	PIWI6KO
RDVJ652	CTRLKO_1_1	1	yes	clone1	control_1	control
RDVJ653	CTRLKO_1_2	2	yes	clone1	control_1	control
RDVJ654	CTRLKO_1_3	3	yes	clone1	control_1	control
RDVJ655	CTRLKO_6_1	1	yes	clone6	control_6	control
RDVJ656	CTRLKO_6_2	2	yes	clone6	control_6	control
RDVJ657	CTRLKO_6_3	3	yes	clone6	control_6	control
RDVJ658	WT_1	1	no	clone0	WT	WT
RDVJ659	WT_2	2	no	clone0	WT	WT
RDVJ660	WT_3	3	no	clone0	WT	WT

```{r}
#Metadata
metadata <- read_delim("Metadata/Metadata3.csv", delim = ";",col_names = TRUE,col_types = c(col_character(),col_character(),col_double(),col_character(),col_double(),col_character(),col_character()))%>% arrange(Library)
```

Condition as the design, and take WT as the baseline. 
```{r}
SampleTable <- metadata %>%
  mutate(Condition= as.factor(Condition)) %>%
  arrange(Library) %>%
  as.data.frame() %>%
  mutate(Condition = fct_relevel(Condition, "WT"))
         
SampleNames = c("PIWI5KO_O_1", "PIWI5KO_O_2", "PIWI5KO_O_3", "PIWI5KO_A_1", "PIWI5KO_A_2", "PIWI5KO_A_3","PIWI6KO_10_1", "PIWI6KO_10_2", "PIWI6KO_10_3", "PIWI6KO_52_1", "PIWI6KO_52_2", "PIWI6KO_52_3","CTRLKO_1_1", "CTRLKO_1_2", "CTRLKO_1_3","CTRLKO_6_1", "CTRLKO_6_2", "CTRLKO_6_3", "WT_1","WT_2","WT_3")

Matrix_TEs = read_delim("Metadata/TEsLau_transcript",
                        delim = "\t",
                        col_names = FALSE)     
```

Takes Salmon quantification data as input
```{r}
#Salmon quantification data

Dir  = "CountDataTE"
FilePath = dir("CountDataTE/", ".sf")
Files=file.path(Dir, FilePath)

names(Files) = paste0(SampleNames)

QuantFiles = tximport(Files, type = "salmon",  tx2gene = Matrix_TEs)
```

General Settings
```{r}
#p value for DESeq2 analysis
pvalue <- 0.01
#log2 fold change for DESeq2 analysis
lfc <- 1
#minimal count of TE to be taken along in statistical analysis
lowCounts <- 10
#FDR
FDR <- 0.05
```


```{r}
#Create a DESeqDataSet
ddsTE = DESeqDataSetFromTximport(QuantFiles,SampleTable,~Condition)
colnames(ddsTE)
#Set the reference as control
ddsTE$Condition <- relevel(ddsTE$Condition, ref = "WT")


#Check PCA plots
rld <- rlog(ddsTE, blind =FALSE)
PCAcondition <- plotPCA(rld, intgroup= "Condition")
print(PCAcondition)

PCAgenotype_condition <- plotPCA(rld, intgroup= c("Genotype","Condition"))
print(PCAgenotype_condition)


#filter low counts
ddsTE = ddsTE[rowSums(counts(ddsTE)) > lowCounts, ]

#run DEseq
DEseq_TE = DESeq(ddsTE,quiet = TRUE)
resultsNames(DEseq_TE)
```

Condition as the design, and take WT as the baseline. 
```{r}
#Piwi6KO vs WT 
results_TE_Piwi6vsWT <- results(DEseq_TE,alpha= pvalue,contrast=c("Condition", "PIWI6KO", "WT"))
plotMA(results_TE_Piwi6vsWT,ylim=c(-10,10))
summary(results_TE_Piwi6vsWT)
results_TE_Piwi6vsWT <- results_TE_Piwi6vsWT %>% as.data.frame() %>% rownames_to_column(var = "TE") %>% as_tibble()
#TEs significantly upregulated at least 2-fold
filter(results_TE_Piwi6vsWT, padj <= pvalue & log2FoldChange >= lfc)


counts_TEs_Piwi6vsWT = counts(DEseq_TE,normalized=TRUE,replaced=FALSE) %>%
 as.data.frame() %>%
 rownames_to_column(var = "TE") %>%
 as_tibble %>%
 group_by(TE) %>%
summarize(meanWT=mean(c(WT_1,WT_2,WT_3)),
 meanPIWI5KO=mean(c(PIWI5KO_O_1,PIWI5KO_O_2,PIWI5KO_O_3,PIWI5KO_A_1,PIWI5KO_A_2,PIWI5KO_A_3)),
 meanPIWI6KO=mean(c(PIWI6KO_10_1,PIWI6KO_10_2,PIWI6KO_10_3,PIWI6KO_52_1,PIWI6KO_52_2,PIWI6KO_52_3)),
 meanCTRLKO=mean(c(CTRLKO_1_1,CTRLKO_1_2,CTRLKO_1_3,CTRLKO_6_1,CTRLKO_6_2,CTRLKO_6_3)))%>%
 mutate(meanWT= meanWT + 0.5,
 meanPIWI5KO= meanPIWI5KO + 0.5,
 meanPIWI6KO= meanPIWI6KO + 0.5,
 meanCTRLKO=meanCTRLKO + 0.5) %>%
 left_join(results_TE_Piwi6vsWT, by = "TE") %>%
 mutate(padj = replace_na(padj, 1),
 sign = ifelse(padj <= FDR, "significant", "NS"), 
 TEtypeSign = str_extract(TE, "(?<=\\d@)\\w+"))
 #TEtypeSign = fct_relevel(TEtypeSign, c("LTR_Gypsy", "DNA_hAT", "RC_Helitron", "Unknown")))

write.table(counts_TEs_Piwi6vsWT, file = "counts_TEs_Piwi6vsWT_withWTbase.tsv", row.names=FALSE, sep="\t")

```

```{r}
#Figure S3E
plotPiwi6vsWT_asbase_colors <- ggplot()+ geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "NS", ], aes(meanWT, meanPIWI6KO),size = 6,alpha=0.6,colour = "gray")+ geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR") , ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#0A8974")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "Unknown" , ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#C2C8E7")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I") , ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E27520")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator") , ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E49C23")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "RC_Helitron", ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E3072A")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "Satellite", ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#6362A9")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "SINE_tRNA", ], aes(meanWT, meanPIWI6KO, colour= TEtypeSign),size = 6,alpha=0.6,colour= "#5FAECC")+
    scale_x_continuous(trans="log2", limits=c(0.5,NA)) + scale_y_continuous(trans="log2", limits=c(0.5,NA)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_abline(slope = 1, intercept= 1) +
    geom_abline(slope = 1, intercept= -1) +
theme(legend.title = element_blank(), panel.background = element_blank(), axis.line = element_line(),legend.key = element_blank()) +theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())

png(file="plotPiwi6vsWT_asbase_colors_TE.png",
   width=484, height=484)
plotPiwi6vsWT_asbase_colors
dev.off()


#Figure S3F
volcanoplotPiwi6vsWT_WTasbase_withouttextcolors <- ggplot()+ geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "NS", ], aes(log2FoldChange, -log10(padj)),size = 6,alpha=0.6,colour = "gray")+  
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#0A8974")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "Unknown" , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#C2C8E7")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E27520")+geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign %in% c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator") , ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E49C23")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "RC_Helitron", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#E3072A")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "Satellite", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#6362A9")+
geom_point(data = counts_TEs_Piwi6vsWT[counts_TEs_Piwi6vsWT$sign == "significant" & counts_TEs_Piwi6vsWT$TEtypeSign == "SINE_tRNA", ], aes(log2FoldChange, -log10(padj), colour= TEtypeSign),size = 6,alpha=0.6,colour= "#5FAECC")+xlim(-10,10)+theme(legend.title = element_blank(), panel.background = element_blank(), axis.line = element_line(),legend.key = element_blank())+geom_vline(xintercept=c(-2, 2), col="black",linetype="longdash")+ geom_hline(yintercept=-log10(0.05), col="black",linetype="longdash")+theme(axis.line=element_line(),panel.grid.major=element_line(color="gray97"),panel.background=element_blank())

png(file="volcanoplotPiwi6vsWT_WTasbase_withouttextcolors_TE.png",
   width=484, height=484)
volcanoplotPiwi6vsWT_WTasbase_withouttextcolors
dev.off() 
```

