---
title: "DESeq2 analysis and visualisation"
author: "Nynke van Eijk"
contact: "nynke.vaneijk@radboudumc.nl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load Packages
```{r, results='hide'}
library(tidyverse)
library(here)
library(patchwork)
library(DESeq2)
library(ggrepel)
library(rstatix)
library(ggpubr)
library(nortest)
library(conflicted)
library(openxlsx)
conflicts_prefer(rstatix::filter)
```

# Antibody 1 (RNA pol II)
## Set-up
```{r, results='hide'}
# Sample list
samples_pol2 <- list(
  "CnT_pol2_wt-1"      = c("condition" = "WT",   "rep" = "1"),
  "CnT_pol2_wt-2"      = c("condition" = "WT",   "rep" = "2"),
  "CnT_pol2_ko10-1"    = c("condition" = "ko10", "rep" = "1"),
  "CnT_pol2_ko10-2"    = c("condition" = "ko10", "rep" = "2"),
  "CnT_pol2_ko52-1"    = c("condition" = "ko52", "rep" = "1"),
  "CnT_pol2_ko52-2"    = c("condition" = "ko52", "rep" = "2"))

# Column data
## condition  = WT, ko10, ko52 (three groups)
## condition2 = WT, KO         (ko10 and ko52 pooled)
coldata_pol2 <- map_dfr(names(samples_pol2), function(sp) {
  data.frame(sample    = sp,
             condition = samples_pol2[[sp]][["condition"]],
             rep       = samples_pol2[[sp]][["rep"]])}) %>%
  mutate(condition  = factor(condition, levels = c("WT", "ko10", "ko52")),
         condition2 = factor(ifelse(condition == "WT", "WT", "KO"), levels = c("WT", "KO"))) %>%
  column_to_rownames("sample")

# Count matrix
## Reads the ${sample}_counts.tsv files produced by 4_counting.sh
count_matrix_pol2 <- map_dfc(names(samples_pol2), function(sp) {
  read_tsv(paste0(sp, "_counts.tsv"),
           col_names = c("transposon", "counts")) %>%
    { if (sp == names(samples_pol2)[1]) . else select(., counts) } %>%
    rename_with(~ sp, "counts")}) %>%
  column_to_rownames("transposon") %>%
  as.matrix() %>%
  round() %>%
  .[rowSums(.) > 0, ]
```

## DESeq2 analysis
```{r, results='hide'}
# Create a DESeq2 DataSet to see the difference between WT, ko10 and ko52
dds_pol2 <- DESeqDataSetFromMatrix(countData = count_matrix_pol2,
                                   colData   = coldata_pol2,
                                   design    = ~condition)
# Create a DESeq2 DataSet to see the difference between WT and KO (ko10 and ko52 together)
dds2_pol2 <- DESeqDataSetFromMatrix(countData = count_matrix_pol2,
                                    colData   = coldata_pol2,
                                    design    = ~condition2)

# DESeq2
## DESeq_pol2  is used for the normalised counts in the violin plots
dds_pol2$condition <- relevel(dds_pol2$condition, ref = "WT")
DESeq_pol2 <- DESeq(dds_pol2, quiet = TRUE)
## DESeq2_pol2 is used for the KO vs WT statistics in the volcano plots
dds2_pol2$condition2 <- relevel(dds2_pol2$condition2, ref = "WT")
DESeq2_pol2 <- DESeq(dds2_pol2, quiet = TRUE)
```

# Antibody 2 (H3K9me3)
## Set-up
```{r, results='hide'}
# Sample list
samples_H3K9me3 <- list(
  "CnT_H3K9me3_wt-1"   = c("condition" = "WT",   "rep" = "1"),
  "CnT_H3K9me3_wt-2"   = c("condition" = "WT",   "rep" = "2"),
  "CnT_H3K9me3_ko10-1" = c("condition" = "ko10", "rep" = "1"),
  "CnT_H3K9me3_ko10-2" = c("condition" = "ko10", "rep" = "2"),
  "CnT_H3K9me3_ko52-1" = c("condition" = "ko52", "rep" = "1"),
  "CnT_H3K9me3_ko52-2" = c("condition" = "ko52", "rep" = "2"))

# Column data
## condition  = WT, ko10, ko52 (three groups)
## condition2 = WT, KO         (ko10 and ko52 pooled)
coldata_H3K9me3 <- map_dfr(names(samples_H3K9me3), function(sh) {
  data.frame(sample    = sh,
             condition = samples_H3K9me3[[sh]][["condition"]],
             rep       = samples_H3K9me3[[sh]][["rep"]])}) %>%
  mutate(condition  = factor(condition, levels = c("WT", "ko10", "ko52")),
         condition2 = factor(ifelse(condition == "WT", "WT", "KO"), levels = c("WT", "KO"))) %>%
  column_to_rownames("sample")

# Count matrix
## Reads the ${sample}_counts.tsv files produced by 4_counting.sh
count_matrix_H3K9me3 <- map_dfc(names(samples_H3K9me3), function(sh) {
  read_tsv(paste0(sh, "_counts.tsv"),
           col_names = c("transposon", "counts")) %>%
    { if (sh == names(samples_H3K9me3)[1]) . else select(., counts) } %>%
    rename_with(~ sh, "counts")}) %>%
  column_to_rownames("transposon") %>%
  as.matrix() %>%
  round() %>%
  .[rowSums(.) > 0, ]
```

## DESeq2 analysis
```{r, results='hide'}
# Create a DESeq2 DataSet to see the difference between WT, ko10 and ko52
dds_H3K9me3 <- DESeqDataSetFromMatrix(countData = count_matrix_H3K9me3,
                                      colData   = coldata_H3K9me3,
                                      design    = ~condition)
# Create a DESeq2 DataSet to see the difference between WT and KO (ko10 and ko52 together)
dds2_H3K9me3 <- DESeqDataSetFromMatrix(countData = count_matrix_H3K9me3,
                                       colData   = coldata_H3K9me3,
                                       design    = ~condition2)

# DESeq2
## DESeq_H3K9me3  is used for the normalised counts in the violin plots
dds_H3K9me3$condition <- relevel(dds_H3K9me3$condition, ref = "WT")
DESeq_H3K9me3 <- DESeq(dds_H3K9me3, quiet = TRUE)
## DESeq2_H3K9me3 is used for the KO vs WT statistics in the volcano plots
dds2_H3K9me3$condition2 <- relevel(dds2_H3K9me3$condition2, ref = "WT")
DESeq2_H3K9me3 <- DESeq(dds2_H3K9me3, quiet = TRUE)
```

# Violin plots
## Set-up
```{r, results='hide'}
# Standard colors, for the outline of the violins
base_colors <- c("WT" = "grey", "ko10" = "#0A8974", "ko52" = "#5FAECC")

# Lighter variant colors, used to fill the violins
light_colors <- sapply(base_colors, function(hex) {
  rgb_vals <- col2rgb(hex) / 255
  light <- rgb_vals + (1 - rgb_vals) * 0.6
  rgb(light[1], light[2], light[3])
})

# Shared theme for the violin and dot panels
panel_theme <- list(theme_bw(),
                    theme(legend.title = element_blank(), 
                          legend.key = element_blank(), 
                          legend.position = "none", 
                          panel.background= element_rect(fill = "white", color = NA), 
                          axis.line=element_line()))

# Shared y-axis for the violin and dot panels
count_scale <- list(scale_y_continuous(trans="log2", limits = c(1,1000000)), 
                    ylab("Mean counts + 0.5 (log2)"))

# Shared aesthetic for the violin and dot panels
panel_aes <- aes(x = condition, y = counts, fill = condition, color = condition)

# Shared layers for the violin panels
violin_layers <- c(list(geom_violin(linewidth=0.8, alpha = 0.5), 
                        scale_fill_manual(values = light_colors), 
                        scale_color_manual(values = base_colors)), 
                        count_scale, panel_theme)

# Extra layers for the paired violin panels (p3 & p4)
paired_layers <- list(geom_line(aes(group = transposon), color = "grey40", linewidth = 0.4, alpha = 0.3),
                      geom_point(size = 1.5, position = position_dodge(0), color = "black"),
                      scale_size_identity())

# Significance brackets for the violin panels
pvalues <- function(pwc) stat_pvalue_manual(pwc, label = "p.adj.signif", tip.length = 0.01, step.increase = 0.08)

# Long format of the condition means, optionally restricted to a set of transposons
mean_counts_long <- function(norm_counts, transposon_filter = NULL) {
  norm_counts %>%
    { if (!is.null(transposon_filter)) filter(., transposon %in% transposon_filter) else . } %>%
    select(transposon, meanWT, meanko10, meanko52) %>%
    pivot_longer(-transposon, names_to = "condition", values_to = "counts") %>%
    mutate(condition = factor(condition, 
                              levels = c("meanWT", "meanko10", "meanko52"), 
                              labels = c("WT", "ko10", "ko52")))}

# Long format of the individual replicates for one transposon
replicate_counts_long <- function(norm_counts, transposon_id) {
  norm_counts %>%
    filter(grepl(transposon_id, transposon)) %>%
    select(transposon, matches("-[12]$")) %>%
    pivot_longer(-transposon, names_to = "sample", values_to = "counts") %>%
    mutate(condition = factor(case_when(grepl("wt", sample) ~ "WT",
                                        grepl("ko10", sample) ~"ko10",
                                        grepl("ko52", sample) ~"ko52"),
                                        levels = c("WT", "ko10", "ko52")))}
```

## Extracting normalised counts
```{r, results='hide'}
# RNA pol II
norm_counts_pol2 <- counts(DESeq_pol2, normalized = TRUE, replaced = FALSE) %>%
  as.data.frame() %>%
  rownames_to_column("transposon") %>%
  mutate(meanWT   = (`CnT_pol2_wt-1` + `CnT_pol2_wt-2`) / 2 + 0.5,
         meanko10 = (`CnT_pol2_ko10-1` + `CnT_pol2_ko10-2`) / 2 + 0.5,
         meanko52 = (`CnT_pol2_ko52-1` + `CnT_pol2_ko52-2`) / 2 + 0.5)

# H3K9me3
norm_counts_H3K9me3 <- counts(DESeq_H3K9me3, normalized = TRUE, replaced = FALSE) %>%
  as.data.frame() %>%
  rownames_to_column("transposon") %>%
  mutate(meanWT   = (`CnT_H3K9me3_wt-1` + `CnT_H3K9me3_wt-2`) / 2 + 0.5,
         meanko10 = (`CnT_H3K9me3_ko10-1` + `CnT_H3K9me3_ko10-2`) / 2 + 0.5,
         meanko52 = (`CnT_H3K9me3_ko52-1` + `CnT_H3K9me3_ko52-2`) / 2 + 0.5)
```

## Statistics
```{r, results='hide'}
# Normality test
## Reports the outliers per condition and tests each condition for normality
run_norm_test <- function(norm_counts, transposon_filter = NULL) {
  df <- mean_counts_long(norm_counts, transposon_filter) %>%
    mutate(counts = log2(counts))
  ## Check for outliers
  outlier_summary <- df %>%
    group_by(condition) %>%
    identify_outliers(counts) %>%
    group_by(condition) %>%
    summarise(n_outliers = n(), contains_AalERV1 = "Consensus_28_AeERV1loci" %in% transposon)
  message("Outliers:"); print(data.frame(outlier_summary))
  ## Test for normality
  normality <- df %>%
    group_by(condition) %>%
    summarise(n = n(),
              test_used = if (n() > 5000) "Anderson-Darling" else "Shapiro-Wilk",
              p_value   = if (n() > 5000) nortest::ad.test(counts)$p.value
                          else shapiro.test(counts)$p.value)
  message("Normality:"); print(data.frame(normality))}

# Repeated measures ANOVA
## Reports the ANOVA table and returns the pairwise paired t-tests
run_paired_stats <- function(norm_counts, transposon_filter = NULL) {
  df <- mean_counts_long(norm_counts, transposon_filter) %>%
    mutate(counts = log2(counts))
  ## ANOVA
  message("ANOVA:")
  print(get_anova_table(anova_test(data = df, dv = counts, wid = transposon, within = condition, type = 3)))
  ## Post-hoc: paired t-tests with Bonferroni correction
  pwc <- df %>%
    pairwise_t_test(counts ~ condition, paired = TRUE, p.adjust.method = "bonferroni") %>%
    add_xy_position(x = "condition")
  return(pwc)}

# Friedman test
## Reports the Friedman test and effect size, and returns the pairwise Wilcoxon tests
run_nonparam_stats <- function(norm_counts, transposon_filter = NULL) {
  df <- mean_counts_long(norm_counts, transposon_filter) %>%
    mutate(counts = log2(counts))
  ## Friedman test (non-parametric equivalent of the repeated measures ANOVA)
  message("Friedman test:")
  print(friedman_test(df, counts ~ condition | transposon))
  ## Effect size
  message("Effect size (Kendall's W):")
  print(friedman_effsize(df, counts ~ condition | transposon))
  ## Post-hoc: Wilcoxon signed-rank with Bonferroni correction
  pwc <- df %>%
    wilcox_test(counts ~ condition, paired = TRUE, p.adjust.method = "bonferroni") %>%
    add_xy_position(x = "condition")
  return(pwc)}

# Note:
## run_norm_test() was run once for every subset below; the results are not part of thisdocument. 
## Based on those results the choice for an ANOVA or a Friedman test was made.
```

## AalERV1
```{r, results='hide'}
# AalERV1 consensus

# RNA pol II
## Dot plot
p1 <- replicate_counts_long(norm_counts_pol2, "Consensus_28_AeERV1loci") %>%
  ggplot(panel_aes) +
  geom_point(size=4.5) +
  scale_color_manual(values = base_colors) +
  count_scale +
#  ggtitle("RNA pol II")+
  xlab("AalERV1 consensus") +
  panel_theme

# H3K9me3
## Dot plot
p2 <- replicate_counts_long(norm_counts_H3K9me3, "Consensus_28_AeERV1loci") %>%
  ggplot(panel_aes) +
  geom_point(size=4.5) +
  scale_color_manual(values = base_colors) +
  count_scale +
#  ggtitle("H3K9me3")+
  xlab("AalERV1 consensus") +
  panel_theme
```

## Up-regulated more than 16-fold in RNA-seq
```{r, results='hide'}
# Up-regulated more than 16-fold in RNA-seq
## No family-3394, ERV1 consensus added
significant_up_more <- read_tsv(here("significantupregulatedmorethan16_RNAseq.tsv"),
                                col_names = c("transposon", "start", "end"))

# RNA pol II
## Statistics: repeated measures ANOVA
pwc_p3 <- run_paired_stats(norm_counts_pol2,    significant_up_more$transposon)
## Violin plot
p3 <- mean_counts_long(norm_counts_pol2, significant_up_more$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
  paired_layers +
#  ggtitle("RNA pol II")+
  xlab("Up-regulated > 16-fold") +
  pvalues(pwc_p3)

# H3K9me3
## Statistics: repeated measures ANOVA
pwc_p4 <- run_paired_stats(norm_counts_H3K9me3, significant_up_more$transposon)
## Violin plot
p4 <- mean_counts_long(norm_counts_H3K9me3, significant_up_more$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
  paired_layers +
#  ggtitle("H3K9me3")+
  xlab("Up-regulated > 16-fold") +
  pvalues(pwc_p4)
```

## All repeat annotations
```{r, results='hide'}
# All repeat annotations

# RNA pol II
## Statistics: Friedman test
pwc_p5 <- run_nonparam_stats(norm_counts_pol2)
## Violin plot
p5 <- mean_counts_long(norm_counts_pol2) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("RNA pol II")+
  xlab("All repeats") +
  pvalues(pwc_p5)

# H3K9me3
## Statistics: Friedman test
pwc_p6 <- run_nonparam_stats(norm_counts_H3K9me3)
## Violin plot
p6 <- mean_counts_long(norm_counts_H3K9me3) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("H3K9me3")+
  xlab("All repeats") +
  pvalues(pwc_p6)
```

## Not significantly up-regulated in RNA-seq
```{r, results='hide'}
# Not significantly up-regulated in RNA-seq
non_significant <- read_tsv(here("nonsignificant_RNAseq.tsv"),
                            col_names = c("transposon", "start", "end"))

# RNA pol II
## Statistics: Friedman test
pwc_p7 <- run_nonparam_stats(norm_counts_pol2,    non_significant$transposon)
## Violin plot
p7 <- mean_counts_long(norm_counts_pol2, non_significant$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("RNA pol II")+
  xlab("Not-significant") +
  pvalues(pwc_p7)

# H3K9me3
## Statistics: Friedman test
pwc_p8 <- run_nonparam_stats(norm_counts_H3K9me3, non_significant$transposon)
## Violin plot
p8 <- mean_counts_long(norm_counts_H3K9me3, non_significant$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("H3K9me3")+
  xlab("Not-significant") +
  pvalues(pwc_p8)
```

## Down-regulated in RNA-seq
```{r, results='hide'}
# Down-regulated in RNA-seq
significant_down <- read_tsv(here("significant_downregulated_RNAseq.tsv"),
                             col_names = c("transposon", "start", "end"))

# RNA pol II
## Statistics: Friedman test
pwc_p9 <- run_nonparam_stats(norm_counts_pol2,    significant_down$transposon)
## Violin plot
p9 <- mean_counts_long(norm_counts_pol2, significant_down$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("RNA pol II")+
  xlab("Down-regulated") +
  pvalues(pwc_p9)

# H3K9me3
## Statistics: Friedman test
pwc_p10 <- run_nonparam_stats(norm_counts_H3K9me3, significant_down$transposon)
## Violin plot
p10 <- mean_counts_long(norm_counts_H3K9me3, significant_down$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("H3K9me3")+
  xlab("Down-regulated") +
  pvalues(pwc_p10)
```

## Up-regulated in RNA-seq
```{r, results='hide'}
# Up-regulated in RNA-seq
## No family-3394, ERV1 consensus added
significant_up <- read_tsv(here("significant_upregulated_RNAseq.tsv"),
                           col_names = c("transposon", "start", "end"))

# RNA pol II
## Statistics: Friedman test
pwc_p11 <- run_nonparam_stats(norm_counts_pol2,    significant_up$transposon)
## Violin plot
p11 <- mean_counts_long(norm_counts_pol2, significant_up$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("RNA pol II")+
  xlab("Up-regulated") +
  pvalues(pwc_p11)

# H3K9me3
## Statistics: Friedman test
pwc_p12 <- run_nonparam_stats(norm_counts_H3K9me3, significant_up$transposon)
## Violin plot
p12 <- mean_counts_long(norm_counts_H3K9me3, significant_up$transposon) %>%
  ggplot(panel_aes) +
  violin_layers +
#  ggtitle("H3K9me3")+
  xlab("Up-regulated") +
  pvalues(pwc_p12)
```

# Volcano/Scatter plots
## Set-up
```{r, results='hide'}
# p-value for DESeq2 analysis
pvalue <- 0.01

# False Discovery Rate (FDR)
FDR <- 0.05

# Transposable element groups
TE_GROUPS <- list(LTR       = c("LTR_Pao","LTR_Gypsy","LTR_Copia","LTR", "AeERV1"),
                  LINE      = c("LINE","LINE_R1","LINE_RTE","LINE_Penelope","LINE_L1","LINE_I"),
                  DNA       = c("DNA","DNA_CMC","DNA_hAT","DNA_Kolobok","DNA_MULE","DNA_PIF","DNA_Sola","DNA_TcMar","DNA_Zator"),
                  RC        = "RC_Helitron",
                  Satellite = "Satellite",
                  SINE      = "SINE_tRNA",
                  Unknown   = "Unknown")

# Colors for the transposable element groups
TE_COLORS <- c(LTR       = "#0A8974",  
               LINE      = "#E27520",  
               DNA       = "#E49C23",  
               RC        = "#E3072A",  
               Satellite = "#6362A9",
               SINE      = "#5FAECC",  
               Unknown   = "#C2C8E7",
               NS        = "gray")

# Shared layers for the scatter panels
ablines   <- list(geom_abline(slope=1, intercept=0),
                  geom_abline(slope=1, intercept=1),
                  geom_abline(slope=1, intercept=-1))
log2scale <- list(scale_x_continuous(trans="log2", limits=c(0.5,NA)),
                  scale_y_continuous(trans="log2", limits=c(0.5,NA)))

# Shared layers for the volcano panels
vlines <- list(geom_vline(xintercept=c(-2, 2), linetype="longdash"),
               geom_hline(yintercept=-log10(0.05), linetype="longdash"))

# Shared theme for the scatter and volcano panels
base_theme <- theme(legend.title = element_blank(), 
                    legend.key = element_blank(),
                    legend.position = "none",
                    panel.grid.major = element_line("gray97"),
                    panel.background= element_rect(fill = "white", color = NA),
                    axis.line=element_line())
```

## Piwi6 KO vs WT results
```{r, results='hide'}
# RNA pol II
## Extract the pvalue, padj, etc.
KOvsWT_pol2 <- results(DESeq2_pol2, alpha = pvalue, contrast = c("condition2", "KO", "WT")) %>%
  as.data.frame() %>%
  rownames_to_column(var = "transposon")
write.xlsx(KOvsWT_pol2, "SupplementaryTable_CnT_Pol2.xlsx")

# H3K9me3
## Extract the pvalue, padj, etc.
KOvsWT_H3K9me3 <- results(DESeq2_H3K9me3, alpha = pvalue, contrast = c("condition2", "KO", "WT")) %>%
  as.data.frame() %>%
  rownames_to_column(var = "transposon")
write.xlsx(KOvsWT_H3K9me3, "SupplementaryTable_CnT_H3K9me3.xlsx")
```

## Annotated counts table
```{r, results='hide'}
# Extract the normalised counts for the requested condition means, join the DESeq2 statistics,
# and annotate with TE type, display name, significance call and colour group
annotate_counts <- function(dds, res, mean_spec) {
  norm <- counts(dds, normalized = TRUE, replaced = FALSE) %>%
    as.data.frame()
  means <- imap(mean_spec, ~ unname(rowMeans(norm[, .x, drop = FALSE])) + 0.5)
  as_tibble(means) %>%
    mutate(transposon = rownames(norm), .before = 1) %>%
    left_join(res, by = "transposon") %>%
    mutate(padj        = replace_na(padj,1),
           sign        = ifelse (padj<= FDR, "significant", "NS"),
           TEtypeSign  = coalesce(str_extract(transposon, "(?<=\\d@)\\w+"),
                                  str_extract(transposon, "(?<=Consensus_\\d{1,3}_)[A-Za-z0-9]+(?=loci)")),
           TE_name     = coalesce(str_replace(str_extract(transposon, "family-\\d+@[^@]+$"), 
                                              "(family-\\d+)@(.+)", "\\2 \\1") %>%
                                  str_replace("_", "/") %>% 
                                  str_replace(" ", "\n"),
                                  str_extract(transposon, "(?<=Consensus_\\d{1,3}_)[A-Za-z0-9]+(?=loci)")),
           color_group = case_when(sign == "NS"                   ~ "NS",
                                   TEtypeSign %in% TE_GROUPS$LTR  ~ "LTR",
                                   TEtypeSign %in% TE_GROUPS$LINE ~ "LINE",
                                   TEtypeSign %in% TE_GROUPS$DNA  ~ "DNA",
                                   TEtypeSign %in% TE_GROUPS$RC   ~ "RC",
                                   TEtypeSign == "Satellite"      ~ "Satellite",
                                   TEtypeSign == "SINE_tRNA"      ~ "SINE",
                                   TEtypeSign == "Unknown"        ~ "Unknown",
                                   TRUE ~ "NS"))}

# Extract normalised counts for KO vs KO and add the pvalue, padj, etc.
## RNA pol II
counts_KOvsKO_pol2 <- annotate_counts(
  DESeq2_pol2, KOvsWT_pol2,
  list(meanko10 = c("CnT_pol2_ko10-1", "CnT_pol2_ko10-2"),
       meanko52 = c("CnT_pol2_ko52-1", "CnT_pol2_ko52-2")))

## H3K9me3
counts_KOvsKO_H3K9me3 <- annotate_counts(
  DESeq2_H3K9me3, KOvsWT_H3K9me3,
  list(meanko10 = c("CnT_H3K9me3_ko10-1", "CnT_H3K9me3_ko10-2"),
       meanko52 = c("CnT_H3K9me3_ko52-1", "CnT_H3K9me3_ko52-2")))

# Extract normalised counts for KO vs WT and add the pvalue, padj, etc.
## RNA pol II
counts_KOvsWT_pol2 <- annotate_counts(
  DESeq2_pol2, KOvsWT_pol2,
  list(meanWT = c("CnT_pol2_wt-1", "CnT_pol2_wt-2"),
       meanko = c("CnT_pol2_ko10-1", "CnT_pol2_ko10-2",
                  "CnT_pol2_ko52-1", "CnT_pol2_ko52-2")))

## H3K9me3
counts_KOvsWT_H3K9me3 <- annotate_counts(
  DESeq2_H3K9me3, KOvsWT_H3K9me3,
  list(meanWT = c("CnT_H3K9me3_wt-1", "CnT_H3K9me3_wt-2"),
       meanko = c("CnT_H3K9me3_ko10-1", "CnT_H3K9me3_ko10-2",
                  "CnT_H3K9me3_ko52-1", "CnT_H3K9me3_ko52-2")))
```

## Scatterplot of KO vs KO
```{r, results='hide'}
# RNA pol II
## Scatter plot
s1 <- ggplot(counts_KOvsKO_pol2, aes(meanko10, meanko52)) + 
  geom_point(data = \(df) df[df$sign == "NS", ], size=6, alpha=0.6, colour = "gray") +
  geom_point(data= \(df) df[df$sign == "significant", ], aes(colour = color_group), size=6, alpha=0.6) +
  scale_color_manual(values = TE_COLORS)+
  ablines + 
  log2scale +
  ggtitle("RNA pol II")+
  labs(x = "Piwi6 KO#10 (mean counts + 0.5)", y = "Piwi6 KO#52 (mean counts + 0.5)") +
  base_theme

# H3K9me3
## Scatter plot
s2 <- ggplot(counts_KOvsKO_H3K9me3, aes(meanko10, meanko52)) + 
  geom_point(data = \(df) df[df$sign == "NS", ], size=6, alpha=0.6, colour = "gray") +
  geom_point(data= \(df) df[df$sign == "significant", ], aes(colour = color_group), size=6, alpha=0.6) +
  scale_color_manual(values = TE_COLORS)+
  ablines + 
  log2scale +
  ggtitle("H3K9me3")+
  labs( x = "Piwi6 KO#10 (mean counts + 0.5)", y = "Piwi6 KO#52 (mean counts + 0.5)") +
  base_theme
```

## Volcano plot of KO vs WT
```{r, results='hide'}
# RNA pol II
## Labels: strongly up-regulated repeats, placed to the right of the panel
label_data <- subset(counts_KOvsWT_pol2, log2FoldChange > 2 & sign == "significant" & -log10(padj) > 20)
## Volcano plot
v1 <- ggplot(counts_KOvsWT_pol2, aes(log2FoldChange, -log10(padj)))+
  geom_point(data = \(df) df[df$sign == "NS", ], colour="gray", size= 6, alpha= 0.6)+
  geom_point(data = \(df) df[df$sign == "significant", ], aes(colour= color_group), size= 6, alpha= 0.6) +
  scale_color_manual(values = TE_COLORS) +
  geom_label_repel(data= label_data, aes(label=TE_name), nudge_x = 2, direction = "y", hjust = 0, xlim=c(max(counts_KOvsWT_pol2$log2FoldChange)+2,Inf),
                   segment.color="grey40", segment.size=0.4, segment.alpha=0.8, fill=alpha(c("white"), 0.6), colour="black", size=5, label.padding = unit(0.5, "lines"),
                   box.padding= unit(0.4, "lines"), point.padding = unit(0.2, "lines"), min.segment.length=0, max.overlaps= Inf)+
  coord_cartesian(xlim=c(-7.5,7.5), clip = "off") +
  vlines + 
  labs( x = "Log2 FC (Piwi6 KO/WT)", y = "-Log10 (adjusted P value)") +
  base_theme +
  ggtitle("RNA pol II")+
  theme(plot.margin = margin(5.5, 100, 5.5, 5.5))

# H3K9me3
## Labels: down-regulated repeats, placed to the left of the panel
label_data2 <- subset(counts_KOvsWT_H3K9me3, log2FoldChange > -5 & log2FoldChange < -2 & sign == "significant" & -log10(padj) > 5)
## Volcano plot
v2 <- ggplot(counts_KOvsWT_H3K9me3, aes(log2FoldChange, -log10(padj)))+
  geom_point(data = \(df) df[df$sign == "NS", ], colour="gray", size= 6, alpha= 0.6)+
  geom_point(data = \(df) df[df$sign == "significant", ], aes(colour= color_group), size= 6, alpha= 0.6) +
  scale_color_manual(values = TE_COLORS) +
  geom_label_repel(data= label_data2, aes(label=TE_name), nudge_x = 0, direction = "y", hjust = 0, xlim=c(-Inf, min(counts_KOvsWT_H3K9me3$log2FoldChange)-5),
                   segment.color="grey40", segment.size=0.4, segment.alpha=0.8, fill=alpha(c("white"), 0.6), colour="black", size=5, label.padding = unit(0.5, "lines"),
                   box.padding= unit(0.4, "lines"), point.padding = unit(0.2, "lines"), min.segment.length=0, max.overlaps= Inf)+
  coord_cartesian(xlim=c(-7.5,7.5), clip = "off") +
  vlines + 
  labs( x = "Log2 FC (Piwi6 KO/WT)", y = "-Log10 (adjusted P value)") +
  base_theme +
  ggtitle("H3K9me3")+
  theme(plot.margin = margin(5.5, 5.5, 5.5, 100))
```
