---
title: "heatmap-TE-counts-GFP-GFPPiwi6"
output: html_document
date: "2026-05-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/phd/Piwi6_revision/small RNA data GFP-IP/GFP-Piwi6 counts")
```

```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
library(wesanderson)
library("RColorBrewer")
library(tidyr)
library(dplyr)
```

small RNA counts in GFP, GFP-Piwi6, GFP-Piwi6-YK mapping to TE annoation from Lau lab was saved in an excel sheet. Counts were normalized to million miRNAs, and the ratios of  GFP-Piwi6/GFP normalized counts were calculated. The TEs with >2 ratio (which was 641 TEs) were saved those in a different .csv file named "nrm_counts_filtered_morethan_2_GFP-Piwi6overGFP". 
```{r}
nrmcounts_TE <- df  <- read_delim(
    "nrm_counts_filtered_morethan_2_GFP-Piwi6overGFP.csv",
    delim = ";",
    col_names = c("TE","GFP", "GFP-Piwi6", "GFP-Piwi6-YK"),
    col_types = cols(
        col_character(),
        col_number(),
        col_number(),
        col_number()
    ),
    locale = locale(decimal_mark = ",")
)
```

all TE counts
```{r}
nrmcounts_long <- nrmcounts_TE %>%
  filter(`GFP-Piwi6` > 1000) %>%   # filter at TE level
  select(TE, GFP, `GFP-Piwi6`) %>%
  pivot_longer(
    cols = c(GFP, `GFP-Piwi6`),
    names_to = "Condition",
    values_to = "Expression"
  )

```

TEs were ordered by TE expression in GFP-Piwi6.
Figure S5A
```{r}
nrmcounts_long <- nrmcounts_TE %>%
  filter(`GFP-Piwi6` > 1000) %>%
  select(TE, GFP, `GFP-Piwi6`) %>%
  arrange(desc(`GFP-Piwi6`)) %>%              # order by Piwi6 signal
  mutate(TE = factor(TE, levels = rev(TE))) %>%  # keep this order in plot
  pivot_longer(
    cols = c(GFP, `GFP-Piwi6`),
    names_to = "Condition",
    values_to = "Expression"
  )

min_val  <- min(nrmcounts_long$Expression, na.rm = TRUE)
max_val  <- max(nrmcounts_long$Expression, na.rm = TRUE)
mean_val <- mean(nrmcounts_long$Expression, na.rm = TRUE)

heatmap_all_more_than_1000_countsinGFPPiwi6 <- ggplot(nrmcounts_long, aes(x = Condition, y = TE, fill = Expression)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#E3072A",
    mid = "white",
    high = "#0A8974",
    midpoint = mean_val,
    limits = c(min_val, max_val)
  ) +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title = element_blank()
  )
heatmap_all_more_than_1000_countsinGFPPiwi6
png(file="heatmap_all_more_than_1000_countsinGFPPiwi6.png",
    width=322, height=484)
heatmap_all_more_than_1000_countsinGFPPiwi6
dev.off()
```

TEs were ordered by TE expression in GFP-Piwi6.
Only taking the top20 of the TEs that are expressed highest in GFP-Piwi6 with TE names.
Figure 3C
```{r}
top20_long <- nrmcounts_TE %>%
  filter(`GFP-Piwi6` > 1000) %>%
  select(TE, GFP, `GFP-Piwi6`) %>%
  arrange(desc(`GFP-Piwi6`)) %>%
  slice_head(n = 20) %>%
  mutate(
    TE_label = TE %>%
      sub("^TE@", "", .) %>%
      sub("^(.*)_family[-_](.*)@(.*)$", "\\3 family-\\2", .) %>%
      sub("_", "/", .),
    TE_label = factor(TE_label, levels = rev(TE_label))
  ) %>%
  pivot_longer(
    cols = c(GFP, `GFP-Piwi6`),
    names_to = "Condition",
    values_to = "Expression"
  )

top20_inGFPPiwi6_heatmap <- ggplot(top20_long, aes(x = Condition, y = TE_label, fill = Expression)) +
  geom_tile() +
  scale_fill_gradient2(
   low = "#E3072A",
    mid = "white",
    high = "#0A8974",
    midpoint = mean_val,
    limits = c(min_val, max_val)
  )  +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.text.y = element_text(size = 7),
    axis.ticks.y = element_blank(),
    axis.title = element_blank()
  )
top20_inGFPPiwi6_heatmap
png(file="top20_inGFPPiwi6_heatmap.png",
    width=242, height=484)
top20_inGFPPiwi6_heatmap
dev.off()

```

